---
title: "Multiple Imputation"
author: "Xiaonan Liu"
date: "2023-04-27"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(yaml)
library(here)
library(tidyr)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)
library(png)
library(dplyr)
library(Rcpp)
library(etm)
library(readr)
library(tibble)
library(Publish)
library(rsample)
library(arsenal)
library(survival)
library(survminer)
library(forestplot)
library(pander)
library(naniar)
library(boot)
library(forcats)
library(mice)
library(lattice)

panderOptions('table.alignment.default', function(df)
    ifelse(sapply(df, is.numeric), 'right', 'left'))
panderOptions('table.split.table', Inf)
panderOptions('big.mark', ",")
panderOptions('keep.trailing.zeros', TRUE)
panderOptions('missing', '')


# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
#options(knitr.kable.NA='', knitr.table.format = "html")
#options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)

knitr::opts_knit$set(root.dir = here::here())

my_controls <- tableby.control(
    test = F,
    total = F,
    #numeric.test = "kwt", cat.test = "chisq",
    numeric.stats = c("meansd", "range", "Nmiss"), 
    cat.stats = c("countpct", "Nmiss"),
    stats.labels = list(
      meansd = "Mean (SD)",
      #medianq1q3 = "Median (IQR)",
      range = "Min - Max",
      Nmiss = "Missing",
      Npct="N (Pct)"
    ),
    digits = 2L
  )


# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(config$functions))

derived_path=config$data$derived

HESonly_dem<-readRDS(here::here(file.path(derived_path,"HESonly_dem.rds")))
HESGP_dem<-readRDS(here::here(file.path(derived_path,"HESGP_dem.rds")))

HESonly_PD<-readRDS(here::here(file.path(derived_path,"HESonly_PD.rds")))
HESGP_PD<-readRDS(here::here(file.path(derived_path,"HESGP_PD.rds")))

HESonly_diab<-readRDS(here::here(file.path(derived_path,"HESonly_diab.rds")))
HESGP_diab<-readRDS(here::here(file.path(derived_path,"HESGP_diab.rds")))

```


# PD 

MI for HES only first. 

```{r}
# Use TEU_ethnicgrp for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","BaC_Sex","TownsendDepInd","TEU_FaH_PD","TEU_Smo_Status","HESDth_PD_time", "HESDth_PD_status")

HESonlyMI=HESonly_PD[,MI_vars]

# Check missingness
HESonlyMI%>%miss_var_summary()%>%filter(n_miss!=0)

missing_var=HESonlyMI%>%miss_var_summary()%>%filter(n_miss!=0)%>%pull(variable)

# Check correlation
corr=mixed_assoc(HESonlyMI)
# See which correlation above 0.6
corr%>%filter(x!=y&assoc>0.6)%>%View()
# none
```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESonlyMI,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
Imp_PD<-mice(HESonlyMI,pred=pred,maxit = 20,seed = 100,m=10)
#test<-mice(HESonlyMI,pred=pred,maxit = 2,seed = 100,m=2)
```

```{r}

## 1. Assess Convergence 
plot(Imp_PD)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESonly_PD,imp1=Imp_PD,missing_var=missing_var)

```

```{r post derivation}
long1 <- complete(Imp_PD,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESonly_PD<-as.mids(comp_long1)

saveRDS(comp_long1_HESonly_PD,file=file.path(config$data$derived,'HESonlyMI_PD_derived.rds'))
```

MI for HES+GP.

```{r}
# Use TEU_ethnicgrp for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","BaC_Sex","TownsendDepInd","TEU_FaH_PD","TEU_Smo_Status","HESDthGP_PD_time","HESDthGP_PD_status")

HESGPMI=HESGP_PD[,MI_vars]

missing_var=HESGPMI%>%miss_var_summary()%>%filter(n_miss!=0)%>%pull(variable)

```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESGPMI,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
ImpGP_PD<-mice(HESGPMI,pred=pred,maxit = 20,seed = 100,m=10)

```

```{r}

## 1. Assess Convergence 
plot(ImpGP_PD)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESGP_PD,imp1=ImpGP_PD,missing_var=missing_var)

```

```{r post derivation}
long1 <- complete(ImpGP_PD,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESGP_PD<-as.mids(comp_long1)

saveRDS(comp_long1_HESGP_PD,file=file.path(config$data$derived,'HESGPMI_PD_derived.rds'))
```

# T2D 

```{r Temp change level name for foresplot T2D}
levels(HESonly_diab$Estw_gestational)<-c("No ","Yes ")
levels(HESonly_diab$TEU_evidenceHTN)<-c("No  ","Yes  ")

levels(HESGP_diab$Estw_gestational)<-c("No ","Yes ")
levels(HESGP_diab$TEU_evidenceHTN)<-c("No  ","Yes  ")

```

## Male 

MI for HES only males.

```{r}
# Use TEU_ethnicgrp and BSM_BMI for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","TownsendDepInd","TEU_FaH_diab","BSM_WaistCirc","BSM_BMI","Dem_InPhA_WHO","TEU_evidenceHTN","HESDth_T2D_time", "HESDth_T2D_status")

HESonlyMI_Males=HESonly_diab%>%filter(BaC_Sex=="Male")%>%select(all_of(MI_vars))

# Check missingness
HESonlyMI_Males%>%miss_var_summary()%>%filter(n_miss!=0)

missing_var=HESonlyMI_Males%>%miss_var_summary()%>%filter(n_miss!=0)%>%pull(variable)

# Check correlation
corr=mixed_assoc(HESonlyMI_Males)
# See which correlation above 0.6
corr%>%filter(x!=y&assoc>0.6)%>%View()
# High correlation between BMI and Waist circ
```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESonlyMI_Males,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
Imp_males<-mice(HESonlyMI_Males,pred=pred,maxit = 20,seed = 100,m=10)

```

```{r checking diagnosis}

## 1. Assess Convergence 
plot(Imp_males)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESonlyMI_Males,imp1=Imp_males,missing_var=missing_var)

```


```{r post derivation}
long1 <- complete(Imp_males,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Dem_BSM_BMIcat=factor(cut(BSM_BMI,
                       breaks = c(0, 25, 30, 200),
                       right = FALSE),levels=c("[0,25)", "[25,30)", "[30,200)"),labels= c("Underweight/Normal", "Overweight", "Obese")),
    
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESonlymales<-as.mids(comp_long1)

saveRDS(comp_long1_HESonlymales,file=file.path(config$data$derived,'HESonlyMI_males_derived.rds'))
```

MI for HES+GP males.

```{r}
# Use TEU_ethnicgrp and BSM_BMI for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","TownsendDepInd","TEU_FaH_diab","BSM_WaistCirc","BSM_BMI","Dem_InPhA_WHO","TEU_evidenceHTN","HESDthGP_T2D_time", "HESDthGP_T2D_status")

HESGPMI_Males=HESGP_diab%>%filter(BaC_Sex=="Male")%>%select(all_of(MI_vars))

```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESGPMI_Males,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
ImpGP_males<-mice(HESGPMI_Males,pred=pred,maxit = 20,seed = 100,m=10)

```

```{r}

## 1. Assess Convergence 
plot(ImpGP_males)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESGPMI_Males,imp1=ImpGP_males,missing_var=missing_var)

```


```{r post derivation}
long1 <- complete(ImpGP_males,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Dem_BSM_BMIcat=factor(cut(BSM_BMI,
                       breaks = c(0, 25, 30, 200),
                       right = FALSE),levels=c("[0,25)", "[25,30)", "[30,200)"),labels= c("Underweight/Normal", "Overweight", "Obese")),
    
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESGPmales<-as.mids(comp_long1)

saveRDS(comp_long1_HESGPmales,file=file.path(config$data$derived,'HESGPMI_males_derived.rds'))
```



## Female

MI for HES only females.

```{r}
# Use TEU_ethnicgrp and BSM_BMI for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","TownsendDepInd","TEU_FaH_diab","BSM_WaistCirc","BSM_BMI","Dem_InPhA_WHO","TEU_evidenceHTN","HESDth_T2D_time", "HESDth_T2D_status","Estw_gestational")

HESonlyMI_Females=HESonly_diab%>%filter(BaC_Sex=="Female")%>%select(all_of(MI_vars))

# Check missingness
HESonlyMI_Females%>%miss_var_summary()%>%filter(n_miss!=0)

missing_var=HESonlyMI_Females%>%miss_var_summary()%>%filter(n_miss!=0)%>%pull(variable)

```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESonlyMI_Females,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
Imp_females<-mice(HESonlyMI_Females,pred=pred,maxit = 20,seed = 100,m=10)

```

```{r checking diagnosis}

## 1. Assess Convergence 
plot(Imp_females)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESonlyMI_Females,imp1=Imp_females,missing_var=missing_var)

```


```{r post derivation}
long1 <- complete(Imp_females,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Dem_BSM_BMIcat=factor(cut(BSM_BMI,
                       breaks = c(0, 25, 30, 200),
                       right = FALSE),levels=c("[0,25)", "[25,30)", "[30,200)"),labels= c("Underweight/Normal", "Overweight", "Obese")),
    
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESonlyfemales<-as.mids(comp_long1)

saveRDS(comp_long1_HESonlyfemales,file=file.path(config$data$derived,'HESonlyMI_females_derived.rds'))
```


MI for HES+GP females.

```{r}
# Use TEU_ethnicgrp and BSM_BMI for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","TownsendDepInd","TEU_FaH_diab","BSM_WaistCirc","BSM_BMI","Dem_InPhA_WHO","TEU_evidenceHTN","HESDthGP_T2D_time", "HESDthGP_T2D_status","Estw_gestational")

HESGPMI_Females=HESGP_diab%>%filter(BaC_Sex=="Female")%>%select(all_of(MI_vars))

```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESGPMI_Females,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
ImpGP_females<-mice(HESGPMI_Females,pred=pred,maxit = 20,seed = 100,m=10)

```

```{r}

## 1. Assess Convergence 
plot(ImpGP_females)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESGPMI_Females,imp1=ImpGP_females,missing_var=missing_var)

```

```{r post derivation}
long1 <- complete(ImpGP_females,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Dem_BSM_BMIcat=factor(cut(BSM_BMI,
                       breaks = c(0, 25, 30, 200),
                       right = FALSE),levels=c("[0,25)", "[25,30)", "[30,200)"),labels= c("Underweight/Normal", "Overweight", "Obese")),
    
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESGPfemales<-as.mids(comp_long1)

saveRDS(comp_long1_HESGPfemales,file=file.path(config$data$derived,'HESGPMI_females_derived.rds'))
```


# Dementia

```{r Temp change level names for foresplot}
# So stupid... need to have unique level names for forestplot to work...
levels(HESonly_dem$TEU_Alc_Freq)[1]<-"Never "
levels(HESonly_dem$Dem_Depression)<-c("No ","Yes ")
levels(HESonly_dem$Dem_diab)<-c("No  ","Yes  ")
levels(HESonly_dem$Dem_Depression)<-c("No   ","Yes   ")
levels(HESonly_dem$Dem_HearingLoss)<-c("No    ","Yes    ")
levels(HESonly_dem$TEU_evidenceHTN)<-c("No     ","Yes     ")
levels(HESonly_dem$Dem_SocialIso)<-c("No      ","Yes      ")


levels(HESGP_dem$TEU_Alc_Freq)[1]<-"Never "
levels(HESGP_dem$Dem_Depression)<-c("No ","Yes ")
levels(HESGP_dem$Dem_diab)<-c("No  ","Yes  ")
levels(HESGP_dem$Dem_Depression)<-c("No   ","Yes   ")
levels(HESGP_dem$Dem_HearingLoss)<-c("No    ","Yes    ")
levels(HESGP_dem$TEU_evidenceHTN)<-c("No     ","Yes     ")
levels(HESGP_dem$Dem_SocialIso)<-c("No      ","Yes      ")

```

MI for HES only first. Note: We need to remove ~5% missing values of APOE carrier before MI  because those people were missing either did not have genetic data available or were outliers for missingness/heterozyg.

```{r}
# Use TEU_ethnicgrp and BSM_BMI for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","BaC_Sex","TownsendDepInd","TEU_FaH_dem","TEU_APOE_e4_carrier","Dem_Edu_Qual","TEU_Alc_Freq","Dem_InPhA_WHO","TEU_Smo_Status","Dem_Depression","Dem_diab","Dem_HearingLoss","TEU_evidenceHTN","Dem_SocialIso","BSM_BMI","HESDth_Dem_time", "HESDth_Dem_status")

HESonlyMI=HESonly_dem[,MI_vars]%>%
  # remove missing status of APOE carrier
  filter(!is.na(TEU_APOE_e4_carrier))

# Check missingness
HESonlyMI%>%miss_var_summary()%>%filter(n_miss!=0)

missing_var=HESonlyMI%>%miss_var_summary()%>%filter(n_miss!=0)%>%pull(variable)

# Check correlation
corr=mixed_assoc(HESonlyMI)
# See which correlation above 0.6
corr%>%filter(x!=y&assoc>0.6)%>%View()
# none
```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESonlyMI,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
Imp_dem<-mice(HESonlyMI,pred=pred,maxit = 20,seed = 100,m=10)

```

```{r}

## 1. Assess Convergence 
plot(Imp_dem)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESonly_dem,imp1=Imp_dem,missing_var=missing_var)

```

```{r post derivation}
long1 <- complete(Imp_dem,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Dem_BSM_BMIcat=factor(cut(BSM_BMI,
                       breaks = c(0, 25, 30, 200),
                       right = FALSE),levels=c("[0,25)", "[25,30)", "[30,200)"),labels= c("Underweight/Normal", "Overweight", "Obese")),
    
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESonly_dem<-as.mids(comp_long1)

saveRDS(comp_long1_HESonly_dem,file=file.path(config$data$derived,'HESonlyMI_dem_derived.rds'))
```


MI for HES+GP. Same note on APOE.

```{r}
# Use TEU_ethnicgrp and BSM_BMI for imputation.

MI_vars=c("ID","TEU_BaC_AgeAtRec","TEU_ethnicgrp","BaC_Sex","TownsendDepInd","TEU_FaH_dem","TEU_APOE_e4_carrier","Dem_Edu_Qual","TEU_Alc_Freq","Dem_InPhA_WHO","TEU_Smo_Status","Dem_Depression","Dem_diab","Dem_HearingLoss","TEU_evidenceHTN","Dem_SocialIso","BSM_BMI","HESDthGP_Dem_time", "HESDthGP_Dem_status")

HESGPMI=HESGP_dem[,MI_vars]%>%
  # remove missing status of APOE carrier
  filter(!is.na(TEU_APOE_e4_carrier))

missing_var=HESGPMI%>%miss_var_summary()%>%filter(n_miss!=0)%>%pull(variable)

```


```{r MI}
# Create Imputations (iter=20, M=10)

# Using default imputation method based on variable type

# Create initial predictor matrix for customization
ini <- mice(HESGPMI,maxit = 0,print=FALSE)
pred <- ini$predictorMatrix

# Remove predictors: ID.
pred[,c("ID")] <- 0

# Actual running MI with the customised prediction matrix
ImpGP_dem<-mice(HESGPMI,pred=pred,maxit = 20,seed = 100,m=10)

```


```{r}

## 1. Assess Convergence 
plot(ImpGP_dem)

## 2. Compare summary stats between observed and imputed (or completed) data
dx_list=MI_compare(MI_data=HESGP_dem,imp1=ImpGP_dem,missing_var=missing_var)

```

```{r post derivation}
long1 <- complete(ImpGP_dem,"long",include = TRUE) # important to put 'include = TRUE'

comp_long1<-long1%>%
  mutate(
    Dem_BSM_BMIcat=factor(cut(BSM_BMI,
                       breaks = c(0, 25, 30, 200),
                       right = FALSE),levels=c("[0,25)", "[25,30)", "[30,200)"),labels= c("Underweight/Normal", "Overweight", "Obese")),
    
    Ethnicity=factor(case_when(TEU_ethnicgrp=="White" ~ "White",
                                     is.na(TEU_ethnicgrp) ~ NA_character_,
                                     TRUE~"Other ethnicity"),levels=c("White","Other ethnicity")))


comp_long1_HESGP_dem<-as.mids(comp_long1)

saveRDS(comp_long1_HESGP_dem,file=file.path(config$data$derived,'HESGPMI_dem_derived.rds'))
```















