---
title: "HES vs GP Venn Diagram"
author: "Xiaonan Liu"
date: "2023-02-14"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(include=TRUE, echo=FALSE)

library(yaml)
library(here)
library(tidyr)
library(DiagrammeR)
library(DiagrammeRsvg)
library(magrittr)
library(svglite)
library(rsvg)
library(png)
library(dplyr)
library(Rcpp)
library(etm)
library(readr)
library(tibble)
library(Publish)
library(rsample)
library(eulerr)
library(ggplot2)
# Specify the markdown format for knitr tables, otherwise they're not compatible with kableExtra
#options(knitr.kable.NA='', knitr.table.format = "html")
#options(kableExtra.auto_format = FALSE)
# Set the root directory to the project directory (otherwise knitr works in the directory of the Rmd file by default)

knitr::opts_knit$set(root.dir = here::here())

# Load the project config file for filepaths etc
config = yaml.load_file(here::here("./config.yml"))
source(here::here(config$functions))

Diab_data<-readRDS(here::here(file.path(config$data$derived,"HESGP_diab.rds")))

PD_data<-readRDS(here::here(file.path(config$data$derived,"HESGP_PD.rds")))

Dem_data<-readRDS(here::here(file.path(config$data$derived,"HESGP_dem.rds")))
```


# T2D

```{r T2D,dpi=300,out.width = "60%",out.height = "60%"}

Diab_list<-HESvsGP(data=Diab_data,outcome_label = "T2D",GP_censoring = "T2D_censordate",
                   HESDth_censoring = "HESDth_T2DM_censordate",GP_eventdate = "PC_T2D_eventdate",HESDth_eventdate = "HESDth_T2D_eventdate")

# Venn digram
Diab_list[1]

ggplot() + aes(Diab_list[[2]])+ 
  geom_histogram(binwidth=1, colour="black", fill="dark grey", boundary = 0)+
  labs(title=paste0("HES/Dth dx date-GP dx date among ",length(Diab_list[[2]])," \n overlaping incident T2D cases before GP censoring"),x="Time difference in years")+
  xlim(-10,15)+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

Diab_list[3]

ggplot() + aes(na.omit(Diab_list[[4]]))+ 
  geom_histogram(binwidth=1, colour="black", fill="dark grey", boundary = 0)+
  labs(title=paste0("HES/Dth dx date-GP dx date among ",length(na.omit(Diab_list[[4]]))," \n overlaping incident T2D cases during full follow-up"),x="Time difference in years")+
  xlim(-10,15)+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))


```

If censor at GP: The median of time difference is `r pretty_dp(median(Diab_list[[2]]),2)` years (IQR= `r pretty_confint(quantile(Diab_list[[2]], prob=.25),quantile(Diab_list[[2]], prob=.75),dp=2)`) and the mean is `r pretty_dp(mean(Diab_list[[2]]),2)` years (SD = `r pretty_dp(sd(Diab_list[[2]]),2)`).

If censor at HES/Dth (i.e. full follow-up): The median of time difference is `r pretty_dp(median(na.omit(Diab_list[[4]])),2)` years (IQR=`r pretty_confint(quantile(na.omit(Diab_list[[4]]), prob=.25),quantile(na.omit(Diab_list[[4]]), prob=.75),dp=2)`) and the mean is `r pretty_dp(mean(na.omit(Diab_list[[4]])),2)` years (SD = `r pretty_dp(sd(na.omit(Diab_list[[4]])),2)`).

```{r cumulative prob T2D,dpi=300,out.width = "60%",out.height = "60%"}

outcome_label="T2D"

timediff_cat<-factor(cut(Diab_list[[4]],breaks = c(-100,seq(0,10,by=1),200), right = TRUE),labels=c(seq(0,10,by=1),">10"))

#Tabulate and turn into data.frame
tab <- as.data.frame(table(timediff_cat))
#Add cumFreq and proportions
tab <- transform(tab, cumFreq = cumsum(Freq))%>%
  mutate(cumProb=cumFreq/length(Diab_list[[4]]))

ggplot(data=tab, aes(x=timediff_cat, y=cumProb, group=1)) +
  geom_line()+
  geom_point()+ 
  scale_y_continuous(labels = scales::percent, limits=c(0,1), breaks = seq(0, 1, by = 0.2))+
  labs(title=paste0("Cumulative probability of HES/Dth dx date-GP dx date \namong ",length(Diab_list[[4]])," participants who had incident GP dx of ",outcome_label),x=paste0("Years since first record of ",outcome_label," in GP"),y=paste0("Cumulative probability of first hospital \n admission with ",outcome_label),caption = "Time difference was computed during the full follow-up")+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

```

Note that the cumulative probability plot does not start from 0% because there were people who had HES dx of T2D before GP dx. Each point shows $\frac{\# (HES dx-GP dx \le year)}{\# GP dx}$.


# Parkinson's Disease

```{r PD,dpi=300,out.width = "60%",out.height = "60%"}

PD_list<-HESvsGP(data=PD_data,outcome_label = "PD",GP_censoring = "PD_censordate",
                   HESDth_censoring = "HESDth_PD_censordate",GP_eventdate = "PC_PD_eventdate",HESDth_eventdate = "HESDth_PD_eventdate")

# Venn digram
PD_list[1]

ggplot() + aes(PD_list[[2]])+ 
  geom_histogram(binwidth=1, colour="black", fill="dark grey", boundary = 0)+
  labs(title=paste0("HES/Dth dx date-GP dx date among ",length(PD_list[[2]])," \n overlaping incident PD cases before GP censoring"),x="Time difference in years")+
  xlim(-10,15)+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

PD_list[3]

ggplot() + aes(na.omit(PD_list[[4]]))+ 
  geom_histogram(binwidth=1, colour="black", fill="dark grey", boundary = 0)+
  labs(title=paste0("HES/Dth dx date-GP dx date among ",length(na.omit(PD_list[[4]]))," \n overlaping incident PD cases during full follow-up"),x="Time difference in years")+
  xlim(-10,15)+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))
```

If censor at GP: The median of time difference is `r pretty_dp(median(PD_list[[2]]),2)` years (IQR=`r pretty_confint(quantile(PD_list[[2]], prob=.25),quantile(PD_list[[2]], prob=.75),dp=2)`) and the mean is `r pretty_dp(mean(PD_list[[2]]),2)` years (SD=`r pretty_dp(sd(PD_list[[2]]),2)`).

If censor at HES/Dth (i.e. full follow-up): The median of time difference is `r pretty_dp(median(na.omit(PD_list[[4]])),2)` years (IQR=`r pretty_confint(quantile(na.omit(PD_list[[4]]), prob=.25),quantile(na.omit(PD_list[[4]]), prob=.75),dp=2)`) and the mean is `r pretty_dp(mean(na.omit(PD_list[[4]])),2)` years (SD = `r pretty_dp(sd(na.omit(PD_list[[4]])),2)`).

```{r cumulative prob PD,dpi=300,out.width = "60%",out.height = "60%"}

outcome_label="PD"

timediff_cat<-factor(cut(PD_list[[4]],breaks = c(-100,seq(0,10,by=1),200), right = TRUE),labels=c(seq(0,10,by=1),">10"))

#Tabulate and turn into data.frame
tab <- as.data.frame(table(timediff_cat))
#Add cumFreq and proportions
tab <- transform(tab, cumFreq = cumsum(Freq))%>%
  mutate(cumProb=cumFreq/length(PD_list[[4]]))

ggplot(data=tab, aes(x=timediff_cat, y=cumProb, group=1)) +
  geom_line()+
  geom_point()+ 
  scale_y_continuous(labels = scales::percent, limits=c(0,1), breaks = seq(0, 1, by = 0.2))+
  labs(title=paste0("Cumulative probability of HES/Dth dx date-GP dx date \namong ",length(PD_list[[4]])," participants who had incident GP dx of ",outcome_label),
       x=paste0("Years since first record of ",outcome_label," in GP"),
       y=paste0("Cumulative probability of first hospital \n admission with ",outcome_label),
       caption = "Time difference was computed during the full follow-up")+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

```

# Dementia

```{r Dementia,dpi=300,out.width = "60%",out.height = "60%"}

Dem_list<-HESvsGP(data=Dem_data,outcome_label = "dementia",GP_censoring = "Dem_censordate",
                   HESDth_censoring = "HESDth_Dem_censordate",GP_eventdate = "PC_Dem_eventdate",HESDth_eventdate = "HESDth_Dem_eventdate")

# Venn digram
Dem_list[1]

ggplot() + aes(Dem_list[[2]])+ 
  geom_histogram(binwidth=1, colour="black", fill="dark grey", boundary = 0)+
  labs(title=paste0("HES/Dth dx date-GP dx date among overlaping \n ",length(Dem_list[[2]])," incident dementia cases before GP censoring"),x="Time difference in years")+
  xlim(-5,15)+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

Dem_list[3]

ggplot() + aes(na.omit(Dem_list[[4]]))+ 
  geom_histogram(binwidth=1, colour="black", fill="dark grey", boundary = 0)+
  labs(title=paste0("HES/Dth dx date-GP dx date among overlaping \n ",length(na.omit(Dem_list[[4]]))," incident dementia cases during full follow-up"),x="Time difference in years")+
  xlim(-5,15)+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

```

If censor at GP: The median of time difference is `r pretty_dp(median(Dem_list[[2]]),2)` years (IQR=`r pretty_confint(quantile(Dem_list[[2]], prob=c(.25)),quantile(Dem_list[[2]], prob=c(.75)),dp=2)`) and the mean is `r pretty_dp(mean(Dem_list[[2]]),2)` years (SD = `r pretty_dp(sd(Dem_list[[2]]),2)`).

If censor at HES/Dth (i.e. full follow-up): The median of time difference is `r pretty_dp(median(na.omit(Dem_list[[4]])),2)` years (IQR=`r pretty_confint(quantile(na.omit(Dem_list[[4]]), prob=.25),quantile(na.omit(Dem_list[[4]]), prob=.75),dp=2)`) and the mean is `r pretty_dp(mean(na.omit(Dem_list[[4]])),2)` years (SD = `r pretty_dp(sd(na.omit(Dem_list[[4]])),2)`).

```{r cumulative prob Dem,dpi=300,out.width = "60%",out.height = "60%"}

outcome_label="dementia"

timediff_cat<-factor(cut(Dem_list[[4]],breaks = c(-100,seq(0,10,by=1),200), right = TRUE),labels=c(seq(0,10,by=1),">10"))

#Tabulate and turn into data.frame
tab <- as.data.frame(table(timediff_cat))
#Add cumFreq and proportions
tab <- transform(tab, cumFreq = cumsum(Freq))%>%
  mutate(cumProb=cumFreq/length(Dem_list[[4]]))

ggplot(data=tab, aes(x=timediff_cat, y=cumProb, group=1)) +
  geom_line()+
  geom_point()+ 
  scale_y_continuous(labels = scales::percent, limits=c(0,1), breaks = seq(0, 1, by = 0.2))+
  labs(title=paste0("Cumulative probability of HES/Dth dx date-GP dx date \namong ",length(Dem_list[[4]])," participants who had incident GP dx of \n",outcome_label),
       x=paste0("Years since first record of ",outcome_label," in GP"),
       y=paste0("Cumulative probability of first hospital \n admission with ",outcome_label),
       caption = "Time difference was computed during the full follow-up")+
  theme(axis.text.x=element_text(size=15),axis.text.y=element_text(size=15),
        axis.text=element_text(size=12),
        axis.title=element_text(size=14,face="bold"),
        plot.title = element_text(size=15,face="bold"))

```


# Manual checking

```{r,eval=FALSE}
# cross check

Diab_data%>%filter(GP_T2DM_censordate>HESDth_T2DM_censordate)%>%nrow()

```

```{r,eval=FALSE}
# Q1: How many HES/Dth cases captured between baseline and end of PC
baseline_endPC=Diab_data%>%filter(!is.na(HESDth_T2D_eventdate) & HESDth_T2D_eventdate<=Admin_PC_CensorDate)%>%nrow()

baseline_endPC

# Q2: Within baseline and end of PC, how many cases had both HES/Dth and PC? 

overlap_data=Diab_data%>%filter(!is.na(HESDth_T2D_eventdate) & HESDth_T2D_eventdate<=Admin_PC_CensorDate &
                !is.na(PC_T2D_eventdate) & PC_T2D_eventdate<=Admin_PC_CensorDate)

nrow(overlap_data)

# percentage
nrow(overlap_data)/baseline_endPC
```
```{r Venn diagram,eval=FALSE}

fit <- euler(c("HES/Dth" = baseline_endPC, "GP" = nrow(Extra_GP)+nrow(overlap_data), "HES/Dth&GP" = nrow(overlap_data)))

venn<-plot(fit,quantities = TRUE,fills = "steelblue3")

```


```{r Histogram,eval=FALSE}
# Create histogram of differences in dx time
diffs<-as.numeric((overlap_data$HESDth_T2D_eventdate-overlap_data$PC_T2D_eventdate)/365.25)

# Plot histogram
hist(diffs,main = paste0("HES/Dth dx date-PC dx date among overlaping ",nrow(overlap_data) ," incident T2D cases"),
     xlab = "Time difference in years")
```


```{r,eval=FALSE}
# Q3: How many extra PC cases not captured in HES/Dth till end of PC?

Extra_GP<-Diab_data%>%filter(!is.na(PC_T2D_eventdate) & PC_T2D_eventdate<=Admin_PC_CensorDate &
                !ID%in%overlap_data$ID)

nrow(Extra_GP)

# Total GP cases
nrow(Extra_GP)+nrow(overlap_data)

# Q4: How many of the extra PC cases ended up in HES/Dth at a later point
Extra_GP%>%filter(!is.na(HESDth_T2D_eventdate) &
                    HESDth_T2D_eventdate>Admin_PC_CensorDate &
                    HESDth_T2D_eventdate<=HESDth_T2DM_censordate)%>%nrow()

```


```{r cumulative prob check, eval=FALSE}
outcome_label = "T2D";GP_censoring = "T2D_censordate";
HESDth_censoring = "HESDth_T2DM_censordate";GP_eventdate = "PC_T2D_eventdate";
HESDth_eventdate = "HESDth_T2D_eventdate"
# Rename columns 
data<-Diab_data%>%
    rename(GP_censoring=!!sym(GP_censoring),
           HESDth_censoring=!!sym(HESDth_censoring),
           GP_eventdate=!!sym(GP_eventdate),
           HESDth_eventdate=!!sym(HESDth_eventdate))


data<-data%>%
  # filter people with a GP dx
  filter(!is.na(GP_eventdate) & GP_eventdate<=GP_censoring)%>%
  # create time diff in years HES dx - GP dx for those with a HES dx before full fu
  mutate(timediff=ifelse(!is.na(HESDth_eventdate) & HESDth_eventdate<=HESDth_censoring,(HESDth_eventdate-GP_eventdate)/365.25,NA))%>%
  # categorise timediff in buckets e.g. (0,1]
  mutate(timediff_cat=factor(cut(timediff,breaks = c(-100,seq(0,14,by=1)), right = TRUE),labels=seq(0,14,by=1)))

```












